`mrpkit`: A reproducible, opinionated, workflow for multilevel regression with post-stratification in R

### 2022-01-31 paper/paper.Rmd
Updated motivation in Intro

thanks: "**Feedback:** Comments on the `r format(Sys.time(), '%d %B %Y')` version of this paper are welcome at: lauren.kennedy729@gmail.com. **Funding statement:** We have no direct financial disclosures. **Conflicts statement:** We have no conflicts to disclose. **Code and data accessibility statement:** Code and data are available at: https://github.com/lauken13/mrpkit. We are unable to share direct access to the datasets used in the 2020 US Presidential example in this paper, however they are freely available and we provide detailed instructions for how to obtain them. **Author contributions:** Kennedy had the original idea. Gabry, Kennedy, and Morris made architectural and other high-level decisions. Gabry and Kennedy wrote most of the code and tests in the package. Amaliah wrote the vignette and contributed tests. Alexander wrote the first draft of this paper. All authors contributed to writing the paper, and approved the final version."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: |
    The R package `mrpkit` implements a structured workflow for Multilevel Regression and Poststratification (MRP). The package helps appropriately set-up the survey and post-stratification datasets and then establish a relationship between different variables in these datasets. A substantial amount of the work to do this is specified by the package's methods, saving time, enhancing reproducibility, and reducing the risk of coding errors. The package natively supports multilevel binomial and Bernoulli models fit with `lme4` and `Stan` (via `brms` and `rstanarm`) and allows for the use of custom modeling functions. Summaries and visualizations of the resulting post-stratified estimates can be conducted using the package's methods. This implementation provides a detailed workflow when conducting MRP that will be useful for researchers and in teaching. Using `mrpkit` proactively addresses many common issues and makes it possible to conduct MRP in an efficient way.

@@

Multilevel regression with post-stratification (MRP) is a statistical approach in which surveys are related to each other using a statistical model. This is useful because known biases in one survey can be adjusted by another in a statistically reasonable way. This enables better use of non-representative surveys, consideration of additional information, and for the propagation of uncertainty. However, it can be difficult to use MRP due to this need to relate two different datasets. This seemingly straight-forward requirement imposes considerable hassle in practice and tends to substantially complicate reproducibility. The R [@citer] package `mrpkit` allows users to conduct multilevel regression with post-stratification in an opinionated and specific way. The package provides a workflow whereby users specify two datasets (a sample dataset and a post-stratification dataset), establish a correspondence between them, and then conduct analysis.

@@

The process of conducting MRP involves considering some survey outcome as being due to a combination of many different cells that are typically demographic or geographic in nature, estimating this relationship using a multilevel regression model and then aggregating cell-level estimates to match the population Figure \@ref(fig:laurenflowchart). The survey of interest would typically be a regular survey, such as a political poll of, say, 1,000 respondents, but it could also be a larger survey, such as the American National Election Studies, or similar. The post-stratification dataset would typically be a larger survey, such as, in the case of the US, the American Community Surveys, or a census. The correspondences need to be established for each possible answer for each question, for instance, the age-groups used in both surveys need to be homogenized. The model would typically be a Bayesian hierarchical model, but in principle any variety could be used.

At its core, MRP is a mapping between a survey object and a population object. It is from this mapping that the power of MRP exists, but it can be difficult to establish this mapping in a reproducible way that fits into a typical applied statistics workflow. Making this implementation easier and more reproducible is important because interest in, and the use of, MRP has considerably increased. It can be difficult even for those experienced with MRP to ensure there are no mistakes in this mapping. The `mrpkit` package creates a well-defined workflow for MRP that represents an important contribution for enhancing the reproducibility of MRP analysis.

Our package complements existing packages such as `survey` [@lumleyone; @lumleytwo; @lumleythree], and `DeclareDesign` [@citeDeclareDesign]. The `survey` package is well-established and allows a user to specify a survey design, such as clustered, stratified, etc, generate summary statistics, and conduct sub-population analysis. The `srvyr` package [@citesrvyr] is a variant of `survey` that adds a syntax similar to `dplyr` [@citedplyr]. The `DeclareDesign` package [@citeDeclareDesign] is a newer package that is based around a grammar of research design. These include: data generation, creating treatment and control groups, and sampling. In general, these packages are focused on designing, implementing, and simulating from, surveys. Ours is focused on what is needed for MRP, and we draw on their packages where possible.

In terms of packages to specifically help with MRP, the most common alternative to this package at the moment is for users to do all aspects of the MRP workflow themselves. This can be difficult. In particular, preparing and matching different levels between surveys can be time consuming and potentially introduce undocumented errors. The package `ccesMRPprep` [@shirospackage] is focused on MRP using the Cooperative Congressional Election Study (CCES). More recently the package `tidymrp` [@jospackage] is being developed, which focuses on conducting analysis in a tidy way [@citetidyverse].

>>>>

thanks: "**Feedback:** Comments on the `r format(Sys.time(), '%d %B %Y')` version of this paper are welcome at: lauren.kennedy729@gmail.com. **Funding statement:** We have no direct financial disclosures. **Conflicts statement:** We have no conflicts to disclose. **Code and data accessibility statement:** Code and data are available at: https://github.com/lauken13/mrpkit. We are unable to share direct access to the datasets used in the 2020 US Presidential example in this paper, however they are freely available and we provide detailed instructions for how to obtain them. **Author contributions:** Kennedy had the original idea. Gabry, Kennedy, and Morris made architectural and other high-level decisions. Gabry and Kennedy wrote most of the code and tests in the package. Amaliah wrote the vignette and contributed tests. Alexander wrote the first draft of this paper. All authors contributed to writing the paper, and approved the final version."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: |
    The R package `mrpkit` implements a structured workflow for Multilevel Regression and Poststratification (MRP). The package helps appropriately set-up the survey and post-stratification datasets and then establish a relationship between different variables in these datasets. A substantial amount of the work to do this is specified by the package's methods, saving time, enhancing reproducibility, and reducing the risk of coding errors. The package natively supports multilevel binomial and Bernoulli models fit with `lme4` and `Stan` (via `brms` and `rstanarm`) and allows for the use of custom modeling functions. Summaries and visualizations of the resulting post-stratified estimates can be created using the package's methods. This implementation provides a detailed workflow when conducting MRP that will be useful for students, educators and researchers. Using `mrpkit` proactively addresses many common issues and makes it possible to conduct MRP in an efficient way.

@@

Multilevel regression and post-stratification (MRP [@Gelman97poststratificationinto, @park2004bayesian, @lax2009should]) is a statistical approach used to make population estimates using survey data. The method is often characterised in two steps. The first uses a multilevel regression to estimate the relationship between some binary outcome variable (that we would like a population estimate for) and demographic groups. The second uses this model to predict the probability of the outcome for each combination of demographic variable. A weighted mean of probabilities using the number of individuals in each demographic combination of weights is taken to create a population estimate.

In this paper and package, we argue that MRP is not a simple step procedure but rather a many stepped procedure that includes a number of implicit steps and procedures. Our hope in proposing the package is that these steps and procedures will become explicit, which will provide the opportunity for both more reproducible and defensible code. In doing so, we are proposing an explicit and opinionated workflow for MRP (see Figure \@ref(fig:laurenflowchart)).

To achieve this, we begin by first conceptualising MRP as a relationship between two surveys. We argue that one of the key implicit steps of MRP is a mapping demographics measurements between a survey and a population. Rarely will these two datasets contain the exact same wording with the exact same demographic categories. Even if they do, the two variables will likely have different names in the survey and population datasets. It is from this mapping that the power of MRP exists, but it can be difficult to establish this mapping in a reproducible way that fits into a typical applied statistics workflow. It can be difficult even for those experienced with MRP to ensure there are no mistakes in this mapping.

<!-- The first is the survey that measures the outcome variable in which surveys are related to each other using a statistical model. This is useful because known biases in one survey can be adjusted by another in a statistically reasonable way. This enables better use of non-representative surveys, consideration of additional information, and for the propagation of uncertainty. However, it can be difficult to use MRP due to this need to relate two different datasets. This seemingly straight-forward requirement imposes considerable hassle in practice and tends to substantially complicate reproducibility. The R [@citer] package `mrpkit` allows users to conduct multilevel regression with post-stratification in an opinionated and specific way. The package provides a workflow whereby users specify two datasets (a sample dataset and a post-stratification dataset), establish a correspondence between them, and then conduct analysis. -->

@@

We also note that in the majority of MRP analyses, the post-stratification matrix is not known but is estimated using another, high quality (large, probability with high response rate) survey such as the American Community Survey (CITE). The reason for this approximation is generally due to the scarcity of information collected at the full population. In addition, as MRP has been applied in non-probability surveys (CITE XBOX) and to estimate small areas in the population (CITE DEEP INTERACTIONS), the number of desired demographics to adjust by has grown. This reinforces the concept that MRP is a mapping between two surveys - one which estimates the population well, and one that includes some variable that we wish to estimate in the population.

<!-- The process of conducting MRP involves considering some survey outcome as being due to a combination of many different cells that are typically demographic or geographic in nature, estimating this relationship using a multilevel regression model and then aggregating cell-level estimates to match the population Figure \@ref(fig:laurenflowchart). The survey of interest would typically be a regular survey, such as a political poll of, say, 1,000 respondents, but it could also be a larger survey, such as the American National Election Studies, or similar. The post-stratification dataset would typically be a larger survey, such as, in the case of the US, the American Community Surveys, or a census. The correspondences need to be established for each possible answer for each question, for instance, the age-groups used in both surveys need to be homogenized. The model would typically be a Bayesian hierarchical model, but in principle any variety could be used. -->

With the proliferation of recent advances to the model component of MRP (such as MRSP [@gao2021improving], SRP [@ornstein2020stacked], BARP [@bisbee2019barp]) it seems likely that multiple models will be fit in a single MRP analysis so that that modelling decisions can be compared by their impact on the population estimate. Many existing packages with an MRP focus (e.g., DGO) tend to implement an MRP workflow for their specific modelling method, but it is not easy to compare modelling methods between then without duplication of code. The least restrictive of these is (autoMrP; @broniecki2021automrp), which uses a stacked method but still does not allow for novel methods to be included/compared easily. Our hope with this package is provide a general scaffolding on which advances in MRP research can be laid upon in future.

In terms of packages to specifically help with MRP, the most common alternative to this package at the moment is for users to do all aspects of the MRP workflow themselves. This can be difficult. In particular, preparing and matching different levels between surveys can be time consuming and potentially introduce undocumented decisions if not errors. Efforts like the `ccesMRPprep` [@shirospackage] package focus on the full MRP including data cleaning using the Cooperative Congressional Election Study (CCES), but this code is specific to the CCCES rather than a general solution. It does however, highlight how important the preproccessing of surveys for MRP really is. 
<!-- More recently the package `tidymrp` [@jospackage] is being developed, which focuses on conducting analysis in a tidy way [@citetidyverse]. -->

Our package complements existing packages such as `survey` [@lumleyone; @lumleytwo; @lumleythree], and `DeclareDesign` [@citeDeclareDesign]. The `survey` package is well-established and allows a user to create and use weights for a given survey design (e.g., clustered, stratified, etc). The `srvyr` package [@citesrvyr] is a variant of `survey` that adds a syntax similar to `dplyr` [@citedplyr]. Our proposed package differs from the survey package because MRP is an alternate method to weights based analysis, and because of the difference in the two methods it is not possible to add an MRP method. The `DeclareDesign` package [@citeDeclareDesign] is a newer package that is based around a grammar of research design. These include: data generation, creating treatment and control groups, and sampling. While useful, DeclareDesign has a greater focus on simulation and understanding different designs.
